---
title: "Car MPG Predictor"
subtitle: "An Interactive Shiny Application for Fuel Efficiency Prediction"
author: "Jacob Robinson"
date: "`r Sys.Date()`"
output: 
  ioslides_presentation:
    widescreen: true
    smaller: true
    css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(plotly)
library(DT)
library(knitr)
```

## Slide 1: Introduction {.smaller}

### Car MPG Predictor: Your Fuel Efficiency Calculator

**What is it?**
- An interactive web application built with R Shiny
- Predicts car fuel efficiency (MPG) based on key characteristics
- Uses machine learning to provide accurate estimates

**Why build this?**
- Help users understand factors affecting fuel efficiency
- Demonstrate practical application of predictive modeling
- Provide an intuitive interface for data exploration


## Slide 2: How It Works {.smaller}

### The Prediction Model

The app uses a **multiple linear regression model** trained on the famous `mtcars` dataset:

```{r model-demo, echo=FALSE}
# Load the data
data <- mtcars
data$transmission <- data$am  # am: 0 = automatic, 1 = manual

# Build the prediction model
model <- lm(mpg ~ wt + cyl + am, data = data)

# Show model summary
summary(model)
```

**Key Features:**
- **Weight**: Car weight in thousands of pounds
- **Cylinders**: Number of engine cylinders (4, 6, or 8)
- **Transmission**: Automatic (0) or Manual (1)


## Slide 3: Interactive Features {.smaller}

### User-Friendly Interface

**Input Controls:**
- Slider for car weight (1.5 - 5.5 thousand lbs)
- Dropdown for cylinder selection
- Radio buttons for transmission type

**Real-time Results:**
- Instant MPG prediction
- 95% confidence intervals
- Interactive visualization

```{r demo-prediction, echo=FALSE}
# Example prediction
new_car <- data.frame(wt = 3.2, cyl = 6, am = 0)
prediction <- predict(model, new_car, interval = "prediction")

cat("Example Prediction:\n")
cat("Weight: 3.2k lbs, 6 cylinders, Automatic\n")
cat("Predicted MPG:", round(prediction[1], 1), "\n")
cat("95% Confidence Interval:", round(prediction[2], 1), "-", round(prediction[3], 1), "MPG")
```



## Slide 4: Data Visualization {.smaller}

### Interactive Plots and Data Exploration

The app includes:

**Scatter Plot:**
- Weight vs MPG relationship
- Color-coded by cylinder count
- User's prediction highlighted

**Data Table:**
- Complete mtcars dataset
- Sortable and searchable
- Easy data exploration

```{r data-viz, echo=FALSE, fig.height=4}
# Create a sample visualization
plot_ly(data = data, 
        x = ~wt, 
        y = ~mpg,
        color = ~factor(cyl),
        colors = c("red", "blue", "green"),
        type = "scatter",
        mode = "markers",
        marker = list(size = 8)) %>%
  layout(title = "Car Weight vs MPG by Cylinders",
         xaxis = list(title = "Weight (1000 lbs)"),
         yaxis = list(title = "Miles Per Gallon"))
```



## Slide 5: Benefits and Applications {.smaller}

### Why Choose This App?

**Educational Value:**
- Learn about predictive modeling
- Understand fuel efficiency factors
- Interactive data science experience

**Technical Excellence:**
- Built with modern R Shiny framework
- Responsive design with Bootstrap themes
- Interactive plots with Plotly
- Professional data tables with DT

**Real-world Applications:**
- Car purchasing decisions
- Environmental impact assessment
- Educational tool for statistics classes

**Try it now!** The app is ready for deployment and can be easily shared via GitHub or RPubs.



## Thank You! {.smaller}

**Technologies Used:**
- R Shiny
- Plotly for interactive visualizations
- DT for data tables
- Bootstrap for responsive design

*Built with ❤️ using R and Shiny* 